---
title: "Homework 1"
author: "Cameron Kay, Lea Frank, and Ashley L. Miller"
date: "1/23/2019"
output:
  html_document:
    highlight: zenburn
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

# set knitr options
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 7.5,
                      fig.height = 4.0)

# load packages
library(rio)
library(here)
library(tidyverse)
library(magrittr)
library(ggforce)

# disable scientific notation
options(scipen = 999)

```

```{r load_data}

# load data (7 files total)

data_assess <- import(here::here("data", "assessments.csv"),
                      setclass = "tibble") %>%
  janitor::clean_names()

data_courses <- import(here::here("data", "courses.csv"),
                       setclass = "tibble") %>% 
  janitor::clean_names()

data_std_assess <- import(here::here("data", "studentAssessment.csv"),
                          setclass = "tibble") %>% 
  janitor::clean_names()

data_std_info <- import(here::here("data", "studentInfo.csv"),
               setclass = "tibble") %>% 
  janitor::clean_names()

data_std_regist <- import(here::here("data", "studentRegistration.csv"),
               setclass = "tibble") %>% 
  janitor::clean_names()

data_std_vle <- import(here::here("data", "studentVle.csv"),
               setclass = "tibble") %>% 
  janitor::clean_names()

data_vle <- import(here::here("data", "vle.csv"),
               setclass = "tibble") %>% 
  characterize() %>%
  janitor::clean_names()

```

```{r identify_keys, include=FALSE}

# identify key in course data:
# key = code_module + code_presentation

data_courses %>% 
  count(code_module, code_presentation) %>% 
  filter(n > 1)

# identify key in assessment data:
# key = id_assessment

data_assess %>% 
  count(id_assessment) %>% 
  filter(n > 1)

# identify key in VLE data:
# key = id_site

data_vle %>% 
  count(id_site) %>% 
  filter(n > 1)

# identify key in student info data:
# key = code_module + code_presentation + id_student

data_std_info %>% 
  count(code_module, code_presentation, id_student) %>% 
  filter(n > 1)

# identify key in student registration file
# key = code_module + code_presentation + id_student

data_std_regist %>% 
  count(code_module, code_presentation, id_student) %>% 
  filter(n > 1)

# identify key in student assessment data:
# key = id_student + id_assessment

data_std_assess %>% 
  count(id_student, id_assessment) %>% 
  filter(n > 1)

####################################
# Identify key in student VLE file:
# NO KEY
####################################

data_std_vle %>% 
  count(code_module, code_presentation, id_student, id_site, date, sum_click) %>% 
  filter(n > 1)

# create surrogate key?
data_std_vle %<>% 
  rowid_to_column()

data_std_vle %>%
  count(rowid) %>%
  filter(n > 1)

```

#### 1) Mutating join/data visualization #1 (Ashley's)

  + **Mutating join**

```{r ash_mutating_join}

ash_data <- left_join(data_std_assess, data_std_info, by = "id_student") #joined by id_student

# identify key
# ash_data %>% 
  # count(id_assessment, id_student, code_module, code_presentation) %>% key
  # filter(n > 1)

ash_data <- ash_data %>%
  group_by(id_student) %>%
  summarize(mean_score = mean(score)) %>%
  left_join(ash_data)

ash_data %<>%
  select(id_student, mean_score, gender:age_band) %>%
  distinct(id_student, .keep_all = TRUE)

```

  + **Resulting data visualization(s)**

```{r ash_data_visualization}

ggplot(ash_data, aes(x = age_band, y = mean_score, fill = age_band)) +
  geom_violin() +
  ggforce::geom_sina(size = 0.5) +
  geom_boxplot(alpha = 0) +
  scale_fill_manual(values = c("#157381", "#4D306E", "#C0A5E1")) +
  facet_wrap(~gender) +
  theme_bw() +
  labs(title = "Figure 1.",
       subtitle = "Assessment score distributions by age group and gender",
       x = "Age Group", 
       y = "Mean Assessment Score",
       caption = "Data from https://analyse.kmi.open.ac.uk/open_dataset") +
  theme(legend.position = "none")

ggplot(ash_data, aes(x = highest_education, y = mean_score, fill = highest_education)) +
  geom_violin() +
  ggforce::geom_sina(size = 0.5) +
  geom_boxplot(alpha = 0) +
  scale_fill_viridis_d() +
  theme_bw() +
  labs(title = "Figure 2.",
       subtitle = "Assessment score distributions by level of education",
       x = "Education Level", 
       y = "Mean Assessment Score",
       caption = "Data from https://analyse.kmi.open.ac.uk/open_dataset") +
  theme(legend.position = "none")

```

#### 2) Mutating join/data visualization #2 (Cam's)

  + **Mutating join**

```{r cam_mutating_join}
# join student information with student vle use
cam_data <- left_join(data_std_info, data_std_vle, by = "id_student")

# drop redundant columns
cam_data %<>%
  select(-contains(".y"))

#summarize students average number of clicks per module
cam_data %<>%
  group_by(id_student) %>% # group by student
  summarize(sum_click_m = mean(sum_click, na.rm = TRUE)) %>% # calculate average number of clicks on VLE materials
  filter(!is.na(sum_click_m)) %>% # remove students who only visited once or who have no imd_band
  left_join(data_std_info) %>% # join demographics with summarized data
  select(id_student, # select only the columns of interest 
         sum_click_m,
         gender,
         region,
         highest_education,
         imd_band,
         age_band) %>%
  filter(imd_band != "", # drop cases with no value for imd band
         age_band != "55<=") %>% # drop cases with insufficient data for age band
  mutate(region = if_else(region != "Scotland" & # mutate regions to sort by countries
                          region != "Ireland" &
                          region != "Wales", 
                          "England", 
                          region),
         imd_band_num = parse_number(imd_band)) # change imd_band from a factor to a numeric
```

  + **Resulting data visualization(s)**
  
```{r cam_data_visualization}
# plot
ggplot(cam_data, aes(x = imd_band_num, y = sum_click_m)) +
  geom_jitter(size = .7, alpha = .7, aes(color = sum_click_m), width = 10, height = 1) +
  geom_smooth(method = "lm", se = FALSE, color = "deeppink2", lwd = 1.5, alpha = .8) +
  theme_bw() +
  theme(strip.background = element_rect(fill = "gray22"),
        strip.text = element_text(colour = "white"),
        legend.position = "none") +
  ylim(0, 10) +
  scale_color_gradient2(low = "cyan", mid = "dodgerblue3", high = "dodgerblue3", midpoint = 8) +
  facet_grid(~region) +
  labs(title = "Figure 3.",
       subtitle = "Relationship between deprivation and amount of engagement with VLE materials",
       x = "Index of Multiple Deprivation", 
       y = "Mean interactions with VLE",
     caption = "Data from https://analyse.kmi.open.ac.uk/open_dataset")
  
```

#### 3) Mutating join/data visualization #3 (Lea's)

What I'm interested in looking at is if there is a relationship between how many times students interacted with the VLE material and how well they did on their final exam, and if this relationship depended on whether or not the student had a disability. In other words, did interacting with the VLE material provide an added benefit to students with a disability compared to students without a disability?

```{r mutatejoin3}

# Filter out those students who passed or failed, ignoring those who withdrew from courses. Not sure what "Distinction" meant so will also exclude those rows. I'm going to use distinct cases of each student, since I'm really only interested in whether or not they have a disability.
std_info_passfail <- data_std_info %>% 
  filter(final_result == "Pass" | final_result == "Fail") %>% 
  distinct(id_student, .keep_all = TRUE) %>% 
  select(-code_module, -code_presentation)

# Merge data_std_assess and data_assess so that I can link the assessment scores to the right code_module and code_presentation. Average scores across the different assessment types
std_assess <- left_join(data_std_assess, data_assess, by = "id_assessment") %>% 
  filter(!is.na(score)) %>% 
  group_by(id_student, code_module, code_presentation, assessment_type) %>% 
  summarize(avg_score = mean(score, na.rm = TRUE))

# Take the average VLE interactions for each student in each course, so average sum_click across all materials in a course. The only thing I still don't understand is why student id is unique when there are students that took multiple courses. Missing data? Or just the nature of the different courses, i.e. did some of the courses not have virtual content?
avg_std_vle <- data_std_vle %>% 
  group_by(code_module, code_presentation, id_student) %>% 
  summarize(avg_clicks = mean(sum_click, na.rm = TRUE))

# Merge the number of clicks with the assessment info
std_assess_clicks <- left_join(std_assess, avg_std_vle, 
                               by = c("id_student","code_module", "code_presentation")) %>% 
  filter(!is.na(avg_clicks))

std_assess_clicks %>% 
  ungroup() %>% 
  count(assessment_type)
# Of course I originally tackled this problem trying to look ONLY at final exam scores just to find out there is too much missing data on click information for people that took final exams. Unless I'm doing this wrong?

# Finally, merge the student info with the assessment/VLE data. Will use inner join because I only want observations with all complete data.
lea_data <- inner_join(std_assess_clicks, std_info_passfail, by = "id_student")

# Ran the figure below and there were definitely outliers, so I decided to use a filtering join to remove outliers for avg_clicks and avg_scores.
high_scores_filter <- lea_data %>% 
  ungroup() %>% 
  filter(avg_score > (mean(avg_score) + 3*sd(avg_score))) # no outliers here

high_clicks_filter <- lea_data %>% 
  ungroup() %>% 
  filter(avg_clicks > 15) # I know this is arbitrary, but there were 90 outliers using a criteria of +3 SD, so I just removed that one person that had over 15 clicks

lea_data <- anti_join(lea_data, high_clicks_filter)

# Let's make sure I actually have enough data to do this. 
lea_data %>% 
  ungroup() %>% 
  count(assessment_type, disability)

```

Random(ish) Question: Will having grouped data affect joins?

```{r lea_dataviz}

lea_data %>% 
  ggplot(aes(x = avg_clicks, y = avg_score, color = disability)) +
  geom_point(alpha = .5, size = .6) +
  geom_smooth(method = "lm", se = TRUE, alpha = .3) +
  facet_wrap(~assessment_type) +
  theme_minimal() +
  scale_color_manual(values = c("lightgrey", "darkslateblue")) +
  labs(title = "Do Virtual Learning Environments (VLE) Benefit Those with Reported Disabilities?",
       y = "Average Assessment Scores",
       x = "VLE Usage \n (average number of clicks)",
       col = "Disability")

```


#### 4) Filter join (Cam)

```{r filtering join}


```
  
#### 5) Reproduce the following figure:

```{r reproduce_figure, fig.height = 7.0}

data <- left_join(data_std_assess, data_std_info, by = "id_student") 

data$score = as.numeric(data$score)

####################################
#Reorder levels of highest education
####################################

#levels(factor(data$highest_education)) 

data$highest_education <- factor(data$highest_education, levels = c("No Formal quals", "Lower Than A Level", "A Level or Equivalent", "HE Qualification", "Post Graduate Qualification"))

###############################
#Reorder levels of final result
###############################

#levels(factor(data$final_result)) 

data$final_result <- factor(data$final_result, levels = c("Fail", "Withdrawn", "Pass", "Distinction"))

#Obtain relevant summary stats

data %<>%
  select(-date_submitted:-is_banked, -imd_band:-disability) %>%
  group_by(highest_education, gender, final_result) %>%
  summarize(mean_score = mean(score, na.rm = TRUE))

#Plotting the data

ggplot(data, aes(x = highest_education, y = mean_score)) +
  geom_col(aes(fill = gender),
           position = "dodge",
           alpha = 0.6) +
  scale_fill_manual(values = c("#10A0AC", "#430E70")) +
  facet_wrap(~final_result, nrow = 4) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Average TMA Scores",
       subtitle = "Results Displayed by Education, Gender, and Final Result Designation",
       x = "Highest Education", 
       y = "Average Score",
       caption = "Data from Kuzilek, Hlosta, & Zdrahal (2017).
       Data from https://analyse.kmi.open.ac.uk/open_dataset",
       fill = "Gender")
```